---
title: "counting-points-in-polygons"
author: "jrpepper"
date: "November 16, 2015"
output: html_document
---

#header 1
##Adding in locations of cities in Ontario
This example shows R code for creating buffers from one set of points (in blue). We also take a separate set of points (in red) and count the number of red points within each blue polygon.

```{r, fig.width=9.5}
library(leaflet)
library(maps)
library(rgeos)
library(sp)

cities <- canada.cities[canada.cities$country.etc == "ON",]
coordinates(cities) <- ~long + lat

leaflet(cities) %>% addTiles() %>% addCircleMarkers(radius=4, stroke=2, fillColor = "Red", color="Red")
```

##Add a point with buffers
Adding in an example point with a large buffer:

```{r, fig.width=9.5}
#make an example point
example_points <- data.frame(lat=numeric(), long= numeric()) 
example_points[1,] <- c(44.19,-80.14)
example_points[2,] <- c(43.69,-81.44)
example_points[3,] <- c(44.981,-81.296)

#add coordinates to make it spatial data
coordinates(example_points) <- ~long + lat

#add a buffer around the point
pointsBuffer <- gBuffer(example_points, width=.5, byid = TRUE)

#make leaflet map
leaflet(cities) %>% addTiles() %>% addCircleMarkers(radius=4, stroke=2, fillColor = "Red", color="Red") %>%
  addMarkers(data=example_points) %>%
  addPolygons(data=pointsBuffer)
```

##Perform spatial analysis:
Add up the number of cites in the buffer add it back to example point
```{r, fig.width=9.5}
#count number of cities in each buffer
count <- over(pointsBuffer, cities, fn=length)
pointsBuffer$city_count <- count$name

#make color ramp
palette <- colorNumeric("YlOrRd", pointsBuffer@data$city_count)

#make leaflet map
leaflet(cities) %>% addTiles() %>% addCircleMarkers(radius=4, stroke=0, opacity=0.2, fillColor = "Red", color="Red") %>%
  addPolygons(data=pointsBuffer, color=~palette(city_count),
              popup=~paste(city_count, "cities in this region.")) %>%
  addLegend(title = "Number of cities in buffer", pal = palette, values= pointsBuffer@data$city_count, opacity = 0.9, position="bottomleft")
```

